---
title: SOC 301 Slides
subtitle: Pacific University
author: Chester Ismay
date: Spring 2017
output: 
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: monokai
      highlightLines: true
    countIncrementalSlides: false
    css: example.css
---

```{r include=FALSE}
#options(servr.daemon = TRUE)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.width=9.5, fig.height=4.5)
theme_set(theme_gray(base_size = 26))
```

layout: true

.footer[[Week 08 - Tuesday](https://ismayc.github.io/soc301_s2017/2017/03/14/week-08---tuesday/)]

---

name: week08t
class: center, middle

# Week 08 - Tuesday

# March 14, 2017
---

# Warmup

### Write the code to produce the following table using ALL 5 of the 5MV functions in the `dplyr` package on the `gapminder` data frame in the `gapminder` R package.  
- Your code will be simplest if you use `|` in your `filter`.
- Notice how the data is sorted.

<br>

```{r echo=FALSE}
library(gapminder)
gapminder %>% 
  dplyr::filter(country %in% c("Belgium", "South Africa") | continent == "Oceania") %>%
  group_by(continent, country) %>% 
  summarize(mean_gdp_all_years = mean(gdpPercap)) %>% 
  mutate(mean_minus_one = mean_gdp_all_years - 1) %>% 
  arrange(mean_gdp_all_years) %>% 
  knitr::kable(format = "html")
```

---

# Check in with your groups

- Ensure all members of your group understand the questions from PS6
  - Finish up the Learning Checks from there as needed
  - We will discuss some LCs from there later today
- Keep working on your Group Projects
  - I will be around to discuss your progress with each group
  - I am planning to ask particular members of your group about what has been done as a group


## Today is a work day

---

```{r old, child='temp_deck.Rmd'}
```