---
date: 2017-02-09T13:14:16-08:00
title: Problem Sets Discussion
---

# Problem Set 2 {#ps2}

Learning Checks LC3.2-LC3.7, LC3.9-LC3.10 from [ModernDive](https://ismayc.github.io/moderndiver-book/3-tidy.html)

## LC3.2

Say the following table are stock prices, how would you make this tidy? 

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
data.frame(
  time = as.Date('2009-01-01') + 0:4,
  x = round(rnorm(5, 0, 1), 3),
  y = round(rnorm(5, 0, 2), 3),
  z = round(rnorm(5, 0, 4), 3)
) %>% 
knitr::kable(align = 'c')
```

**Discussion**:  `x`, `y`, and `z` are not variables by themselves, but rather levels of another variable `stock`.  The numbers inside the table themselves are prices.  Therefore, in its current form, `x` corresponds both to the "name of a stock" and the "price of a stock" which violates the rules of tidy data.  We can transform the data to make it **tidy**:

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyr)
data.frame(
  time = as.Date('2009-01-01') + 0:4,
  x = round(rnorm(5, 0, 1), 3),
  y = round(rnorm(5, 0, 2), 3),
  z = round(rnorm(5, 0, 4), 3)
) %>% 
  gather(key = "stock", value = "price", x:z) %>% 
  knitr::kable()
```

***

## LC3.3

What does any *ONE* row in this `flights` dataset refer to?

- A. Data on an airline 
- B. Data on a flight
- C. Data on an airport
- D. Data on multiple flights

**Discussion**:  The correct answer is B here.  The observational unit is flight.  All of the variables are measurements/characteristics of a given flight.  Each row is an observation of one flight containing data about that flight.

***

## LC3.4

What are some examples in this dataset of **categorical** variables?  What makes them different than **quantitative** variables?

**Discussion**:

***

## LC3.5

What does `int`, `dbl`, and `chr` mean in the output above?  If you need a hint, you might want to run `str(flights)` instead.

**Discussion**:  These correspond to the `class` of each variable in R.  `dbl` is the same thing as `num` and represents numbers with decimal values.  `int` corresponds to integers (whole numbers and negative )

***

## LC3.6

How many different columns are in this dataset?

**Discussion**:

***

## LC3.7

How many different rows are in this dataset?

**Discussion**:

***

## LC3.9

***

## LC3.10
