---
date: 2017-02-09T13:14:16-08:00
title: Problem Sets Discussion
output: 
  blogdown::html_page:
    toc: true
    toc_depth: 1
---

***

Page last updated at `r format(Sys.time(), "%A, %B %d, %Y %X %Z")`

```{r echo=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```


# Problem Set 2 {#ps2-key}

Learning Checks LC3.2-LC3.7, LC3.9-LC3.10 from [ModernDive](https://ismayc.github.io/moderndiver-book/3-tidy.html)

## LC3.2

Say the following table are stock prices, how would you make this tidy? 

```{r echo=FALSE}
library(dplyr)
data.frame(
  time = as.Date('2009-01-01') + 0:4,
  x = round(rnorm(5, 0, 1), 3),
  y = round(rnorm(5, 0, 2), 3),
  z = round(rnorm(5, 0, 4), 3)
) %>% 
knitr::kable()
```

**Discussion**:  `x`, `y`, and `z` are not variables by themselves, but rather levels of another variable `stock`.  The numbers inside the table themselves are prices.  Therefore, in its current form, `x` corresponds both to the "name of a stock" and the "price of a stock" which violates the rules of tidy data.  We can transform the data to make it **tidy**:

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyr)
data.frame(
  time = as.Date('2009-01-01') + 0:4,
  x = round(rnorm(5, 0, 1), 3),
  y = round(rnorm(5, 0, 2), 3),
  z = round(rnorm(5, 0, 4), 3)
) %>% 
  gather(key = "stock", value = "price", x:z) %>% 
  knitr::kable(align = c('l', 'c', 'r'))
```


## LC3.3

What does any *ONE* row in this `flights` dataset refer to?

- A. Data on an airline 
- B. Data on a flight
- C. Data on an airport
- D. Data on multiple flights

**Discussion**:  The correct answer is B here.  The observational unit is flight.  All of the variables are measurements/characteristics of a given flight.  Each row is an observation of one flight containing data about that flight.

## LC3.4

What are some examples in this dataset of **categorical** variables?  What makes them different than **quantitative** variables?

**Discussion**:  The **categorical** variables are `carrier`, `tailnum`, `origin`, and `dest`.  The **quantitative** variables are all of the `<int`> and `<dbl>` class variables:  `year`, `month`, `day`, `dep_time`, `sched_dep_time`, `arr_time`, `sched_arr_time`, `arr_delay`, `flight`, `air_time`, `distance`, `hour`, and `minute`.  Note that `time_hour` is labeled as `<dttm>` corresponding to date-time.  This is a special format in R for handling the oddities of dates and times.  Note also that `dep_time` and other similar variables are not labeled as `<dttm>`, but rather as a number.  Think about what problems this might cause in your analyses.

## LC3.5

What does `int`, `dbl`, and `chr` mean in the output above?  If you need a hint, you might want to run `str(flights)` instead.

**Discussion**:  These correspond to the `class` of each variable in R.  `dbl` is the same thing as `num` and represents numbers with decimal values.  `int` corresponds to integers (whole numbers and their negative opposites (-5, 4, 79, 0, 12342, etc.)).  `chr` corresponds to character values/strings and these are almost always categorical variables.  The output of the similar `str(flights)` is below.

```{r}
library(nycflights13)
str(flights)
```


## LC3.6

How many different columns are in this dataset?

**Discussion**:  

```{r}
library(nycflights13)
glimpse(flights)
```

We see above from running `glimpse(flights)` that the number of variables is `r ncol(flights)`.  In a tidy data frame such as `flights`, the number of columns corresponds directly to the number of variables so the number of columns is `r ncol(flights)`. This can also be obtained using the `ncol` function via `ncol(flights)`.

## LC3.7

How many different rows are in this dataset?

**Discussion**:

We see above from running `glimpse(flights)` that the number of observations is `r format(nrow(flights), big.mark = ",", scientific = FALSE)`.  In a tidy data frame such as `flights`, the number of rows corresponds directly to the number of observations/cases so the number of rows is also `r format(nrow(flights), big.mark = ",", scientific = FALSE)`. This can also be obtained using the `nrow` function via `nrow(flights)`.

## LC3.9

What properties of the observational unit do each of `lat`, `lon`, `alt`, `tz`, `dst`, and `tzone` describe for the `airports` data frame?  Note that you may want to use `?airports` to get more information or go to the reference manual for the `nycflights13` package [here](https://cran.r-project.org/web/packages/nycflights13/nycflights13.pdf).

**Discussion**:  

Using `?airports`, we see that `lat` and `lon` correspond to the latitude and longitude of the location of the airport.  `alt` is the altitude of the airport in feet.  `tz` is the offset of the airport's time zone from Greenwich Mean Time.  `dst` corresponds to "Daylight savings time zone" with values of `"A"` for US DST, `"U"` for unknown, and `"N"` for no DST.  Lastly, `tzone` is the name of the time zone in which the airport is located.


## LC3.10

Provide the names of variables in a data frame with at least three variables in which one of them is an identification variable and the other two are not.

**Discussion**:  Many answers are appropriate here.  The identification variable should correspond to the observational unit of the data frame.  Here is one example:

Student ID  |  Major       | Percentage | Grade 
------------|--------------|:------------:|-------
1000        | Sociology    |  95        |  A
1001        | Biology      |  50        |  F
1002        | History      |  76        |  C



