---
date: 2017-02-09T13:14:16-08:00
title: Problem Sets Discussion
output: 
  blogdown::html_page:
    toc: true
    toc_depth: 1
---

***

Page last updated at `r format(Sys.time(), "%A, %B %d, %Y %X %Z")`

```{r echo=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```


# Problem Set 2 {#ps2-key}

Learning Checks LC3.2-LC3.7, LC3.9-LC3.10 from [ModernDive](https://ismayc.github.io/moderndiver-book/3-tidy.html)

## LC3.2

Say the following table are stock prices, how would you make this tidy? 

```{r echo=FALSE}
library(dplyr)
data_frame(
  time = as.Date('2009-01-01') + 0:4,
  Boeing = paste("$", c(173.55, 172.61, 173.86, 170.77, 174.29), sep = ""),
  Amazon = paste("$", format(round(c(174.90, 171.42, 171.58, 173.89, 170.16), 2), nsmall = 2), sep = ""),
  Google = paste("$", c(174.34, 170.04, 173.65, 174.87, 172.19), sep = "")
) %>%  
knitr::kable()
```

**Discussion**:  `Boeing`, `Amazon`, and `Google` are not variables by themselves, but rather levels of another variable `stock`.  The numbers inside the table themselves are prices.  Therefore, in its current form, `Google` corresponds both to the "name of a stock" and the "price of a stock" which violates the rules of tidy data.  We can transform the data to make it **tidy**:

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyr)
data_frame(
  time = as.Date('2009-01-01') + 0:4,
  Boeing = paste("$", c(173.55, 172.61, 173.86, 170.77, 174.29), sep = ""),
  Amazon = paste("$", format(round(c(174.90, 171.42, 171.58, 173.89, 170.16), 2), nsmall = 2), sep = ""),
  Google = paste("$", c(174.34, 170.04, 173.65, 174.87, 172.19), sep = "")
) %>%
  gather(key = "stock", value = "price", Boeing:Google) %>% 
  knitr::kable(align = c('l', 'c', 'r'))
```


## LC3.3

What does any *ONE* row in this `flights` dataset refer to?

- A. Data on an airline 
- B. Data on a flight
- C. Data on an airport
- D. Data on multiple flights

**Discussion**:  The correct answer is B here.  The observational unit is flight.  All of the variables are measurements/characteristics of a given flight.  Each row is an observation of one flight containing data about that flight.

## LC3.4

What are some examples in this dataset of **categorical** variables?  What makes them different than **quantitative** variables?

**Discussion**:  The **categorical** variables are `carrier`, `tailnum`, `origin`, and `dest`.  The **quantitative** variables are all of the `<int`> and `<dbl>` class variables:  `year`, `month`, `day`, `dep_time`, `sched_dep_time`, `arr_time`, `sched_arr_time`, `arr_delay`, `flight`, `air_time`, `distance`, `hour`, and `minute`.  Note that `time_hour` is labeled as `<dttm>` corresponding to date-time.  This is a special format in R for handling the oddities of dates and times.  Note also that `dep_time` and other similar variables are not labeled as `<dttm>`, but rather as a number.  Think about what problems this might cause in your analyses.

## LC3.5

What does `int`, `dbl`, and `chr` mean in the output above?  If you need a hint, you might want to run `str(flights)` instead.

**Discussion**:  These correspond to the `class` of each variable in R.  `dbl` is the same thing as `num` and represents numbers with decimal values.  `int` corresponds to integers (whole numbers and their negative opposites (-5, 4, 79, 0, 12342, etc.)).  `chr` corresponds to character values/strings and these are almost always categorical variables.  The output of the similar `str(flights)` is below.

```{r}
library(nycflights13)
str(flights)
```


## LC3.6

How many different columns are in this dataset?

**Discussion**:  

```{r}
library(nycflights13)
glimpse(flights)
```

We see above from running `glimpse(flights)` that the number of variables is `r ncol(flights)`.  In a tidy data frame such as `flights`, the number of columns corresponds directly to the number of variables so the number of columns is `r ncol(flights)`. This can also be obtained using the `ncol` function via `ncol(flights)`.

## LC3.7

How many different rows are in this dataset?

**Discussion**:

We see above from running `glimpse(flights)` that the number of observations is `r format(nrow(flights), big.mark = ",", scientific = FALSE)`.  In a tidy data frame such as `flights`, the number of rows corresponds directly to the number of observations/cases so the number of rows is also `r format(nrow(flights), big.mark = ",", scientific = FALSE)`. This can also be obtained using the `nrow` function via `nrow(flights)`.

## LC3.9

What properties of the observational unit do each of `lat`, `lon`, `alt`, `tz`, `dst`, and `tzone` describe for the `airports` data frame?  Note that you may want to use `?airports` to get more information or go to the reference manual for the `nycflights13` package [here](https://cran.r-project.org/web/packages/nycflights13/nycflights13.pdf).

**Discussion**:  

Using `?airports`, we see that `lat` and `lon` correspond to the latitude and longitude of the location of the airport.  `alt` is the altitude of the airport in feet.  `tz` is the offset of the airport's time zone from Greenwich Mean Time.  `dst` corresponds to "Daylight savings time zone" with values of `"A"` for US DST, `"U"` for unknown, and `"N"` for no DST.  Lastly, `tzone` is the name of the time zone in which the airport is located.


## LC3.10

Provide the names of variables in a data frame with at least three variables in which one of them is an identification variable and the other two are not.

**Discussion**:  Many answers are appropriate here.  The identification variable should correspond to the observational unit of the data frame.  Here is one example:

Student ID  |  Major       | Percentage | Grade 
------------|--------------|:------------:|-------
1000        | Sociology    |  95        |  A
1001        | Biology      |  50        |  F
1002        | History      |  76        |  C


***

# Problem Set 3 {#ps3-key}

Learning Checks [LC3.11-LC3.13 and LC3.15](https://ismayc.github.io/moderndiver-book/3-tidy.html) AND [LC4.1-LC4.2, LC4.4, LC4.6-LC4.7, and LC4.10-LC4.13](https://ismayc.github.io/moderndiver-book/4-viz.html)

## LC3.11 

What are common characteristics of "tidy" datasets?

**Discussion**: Tidy data sets tend to be in a narrow, long format since they list all possible values of each combination of the different variables.  They have one observation per row, one variable per column, one observational unit per data frame, and the entries on the inside of the table are values, each representing the intersection of a particular observation and a particular variable.

## LC3.12


What makes "tidy" datasets useful for organizing data?

**Discussion**: They allow you to easily summarize data across a variable or to plot data and show different relationships across other variables.  They are incredibly useful in looking for multivariate relationships (more than two variables) in data.  You'll see more of this as the semester goes along.

## LC3.13


How many variables are presented in the table below?  What does each row correspond to? (**Hint:** You may not be able to answer both of these questions immediately but take your best guess.)

```{r echo=FALSE, message=FALSE}
library(dplyr)
library(knitr)
students <- c(4, 6)
faculty <- c(2, 3)
kable(data_frame("students" = students, "faculty" = faculty))
```

**Discussion**: This is an incredibly difficult problem to figure out.  You can guess that `students` and `faculty` should actually be levels of a variable called something like `type` or `role`.  The numbers inside the table represent counts of some sort, so there may be at least two other variables floating around in there.

## LC3.15 

The actual data presented in LC3.13 is given below in tidy data format:

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
role <- c(rep("student", 10), rep("faculty", 5))
sociology <- c(rep(TRUE, 4), rep(FALSE, 6), rep(TRUE, 2), rep(FALSE, 3))
school_type <- c(rep("Public", 6), rep("Private", 4), rep("Public", 3), rep("Private", 2))
kable(data_frame("role" = role, `Sociology?` = sociology,
  `Type of School` = school_type), align = c('l', 'c', 'l'))
```

- What does each row correspond to?  
- What are the different variables in this data frame?  
- The `Sociology?` variable is known as a logical variable.  What types of values does a logical variable take on?


**Discussion**:

- Each row corresponds to a particular observation (an instance of the observational unit).  In this case, each row corresponds to a particular person, either a faculty member or a student.
- The different variables are `role`, `Sociology?`, and `Type of School`.
- Logical variables take on the values of `TRUE` or `FALSE`.


## LC4.1  

Take a look at both the `flights` and `all_alaska_flights` data frames by running `View(flights)` and `View(all_alaska_flights)` in the console. In what respect do these data frames differ?

**Discussion**:   Recall that you don't have the ability to run `View` in DataCamp and we aren't using RStudio just yet.  You can find interactive versions of both data frames [here](https://ismayc.github.io/Effective-Data-Storytelling-using-the-tidyverse/moderndive.html) though.

The only difference between these data frames is that `flights` contains information about 2013 NYC flights from all carriers and `all_alaska_flights` only contains information about flights using Alaskan Airlines.

## LC4.2


What are some practical reasons why `dep_delay` and `arr_delay` have a positive relationship?

**Discussion**:

Flights that are late in leaving the origin airport will almost surely be late in arriving at the destination airport.  Further, a flight that is delayed 15 minutes on departure will probably be delayed around 15 minutes on arrival.  Pilots are able to make up some time in the air sometimes, but a positive relationship is clear here.  As `dep_delay` increases, `arr_delay` also increases.


###

## LC4.6

Create a new scatter-plot using different variables in the `all_alaska_flights` data frame by modifying the example above.  (Run your code in the sandbox in DataCamp and then write your code on your paper. Discuss what findings you have when looking at the plot. Remember that scatter-plots only work when analyzing the relationship between two **numerical** variables!)


**Discussion**:

There are many potential plots you could create here.  Here is one example:

```{r}
library(nycflights13)
library(ggplot2)
all_alaska_flights <- flights %>% filter(carrier == "AS")
ggplot(data = all_alaska_flights, mapping = aes(x = air_time, y = distance)) +
  geom_point()
```

This might initially look like an error but Alaskan Airlines online flies to Seattle from New York City.  Therefore, all of the `distance` values are 2402.  We can see that there was quite a bit of variability in terms of `air_time` though.

<!-- -->

## LC4.7

Why is setting the `alpha` argument value useful with scatter-plots? What further information does it give you that a regular scatter-plot cannot?


**Discussion**:

The `alpha` argument allows for the setting of the transparency of the points.  This allows you to see when many points are over-lapping and to identify over-plotting in your scatter-plot.

## LC4.10


The weather data is recorded hourly. Why does the `time_hour` variable correctly identify the hour of the measurement whereas the `hour` variable does not?

**Discussion**:

The `time_hour` variable gives the precise hour on a given day.  The `hour` variable is just an integer value from 0 to 23 so without other information, `hour` only gives the hour of a particular day.

## LC4.11


Why should line-graphs be avoided when there is not a clear ordering of the horizontal axis?

**Discussion**:

Line-graphs only show trends in the data and are best used with time as the explanatory variable on the horizontal axis.  When there is no clear ordering of the horizontal axis, one could deceive the reader into thinking a pattern/trend exists when, in fact, one does not.

## LC4.12


Why are line-graphs frequently used when time is the explanatory variable?

**Discussion**:

Time on the horizontal axis and a single value for each time point provides a simple way to identify a trend in the variable.  It allows for one to easily answer questions such as "Is the response variable increasing over time?", "Where are there peaks and valleys in the response variable?", etc.

## LC4.13

Plot a time series of a variable other than `temp` for Newark Airport in the first 15 days of January 2013.  (Run your code in the sandbox in DataCamp and then write your code on your paper. Discuss what findings you have when looking at the plot. Remember that scatter-plots only work when analyzing the relationship between two **numerical** variables!)

**Discussion**:

There are many potential answers for this question.  Here is one answer:

```{r}
library(nycflights13)
library(ggplot2)
early_january_weather <- weather %>% 
  filter(origin == "EWR" & month == 1 & day <= 15)
ggplot(data = early_january_weather, mapping = aes(x = time_hour, y = wind_gust)) +
  geom_line()
```

What immediately stands out in this plot are the days with no wind gusts at all and values of 0.  There appears to be only one wind gust over 25 mpg around January 2nd or 3rd.

***

# Problem Set 4 {#ps4-key}

Learning Checks [LC4.14, LC4.16, LC4.18, LC4.19, LC4.21-LC4.28, LC4.30, LC4.33-LC4.34, LC4.36-LC4.37](https://ismayc.github.io/moderndiver-book/4-viz.html)

## LC4.14 

What does changing the number of bins from 30 to 60 tell us about the distribution of temperatures?

**Discussion**: 

Increasing the number of bins allows for us to see the distribution of temperatures in finer detail.

## LC4.16

What would you guess is the "center" value in this distribution? Why did you make that choice?

**Discussion**: 

There is no exact answer here at this point in the course.  I wanted you to think about what the balancing point would be having half the values before it and half after it.  This falls somewhere between 50 and 55.

## LC4.18

What other things do you notice about the faceted plot above? How does a faceted plot help us see relationships between two variables?

**Discussion**: 

The faceted plot allows us to better see the range of temperatures for each month.  In general, a faceted plot can be helpful in seeing trends in one variable compared to another.

## LC4.19

What do the numbers 1-12 correspond to in the plot above? What about 25, 50, 75, 100?

**Discussion**: 

The numbers 1-12 correspond to months in the year 2013.  The nubmers 25, 50, 75, and 100 correspond to Fahrenheit temperatures.

## LC4.21

Does the `temp` variable in the `weather` data set have a lot of variability? Why do you say that?

**Discussion**: 

New York City does have a broad range of temperatures throughout the year.  We see that summer months tend to be much warmer than winter months, so, yes, `temp` in `weather` does have a lot of variability.

## LC4.22

What does the dot at the bottom of the plot for May correspond to? Explain what might have occurred in May to produce this point.

**Discussion**: 

The dot at the bottom of the plot for May corresponds to a low outlier.  This could have been a very cold hour, but since no other hours have that cold of a temperature it is more than likely a measurement error.

## LC4.23

Which months have the highest variability in temperature? What reasons do you think this is?

**Discussion**: 

The spring and fall months have the highest variability in temperature.  These months tend to have cold mornings and warmer high temperatures.

## LC4.24

We looked at the distribution of a continuous variable over a categorical variable here with this boxplot. Why can’t we look at the distribution of one continuous variable over the distribution of another continuous variable? Say, temperature across pressure, for example?

**Discussion**:

A boxplot works well because you can look at the distribution of the continuous variable across the different (often only a few) levels of the categorical variable.  A continuous explanatory variable will have many, many levels making the boxplot very hard to read and interpret.

## LC4.25

Boxplots provide a simple way to identify outliers. Why may outliers be easier to identify when looking at a boxplot instead of a faceted histogram?

**Discussion**: 

A faceted histogram shows the height of the bin and if outliers are all alone they will appear as having a bin of height close to zero, which is almost indistinguishable from zero.  Therefore, they are more likely to be missed when looking at a faceted histogram.  A boxplot will clearly show them.

## LC4.26

Why are histograms inappropriate for visualizing categorical variables?

**Discussion**:

Histograms show how many values fall into different bins of a numerical variable that falls on a continuum.  (This is one reason why the bins are touching.)  Since a categorical variable does not fall on a continuum, histograms are inappropriate.

## LC4.27

What is the difference between histograms and barplots?

**Discussion**: 

The bins touch on a histogram and the bars/bins on a barplot do not.  Histograms show the distribution of a continuous variable whereas barplots show the distribution of a categorical variable.

## LC4.28

How many Envoy Air flights departed NYC in 2013?

**Discussion**: 

Since Envoy Air corresponds to `carrier` code `MQ` via the `airports` data frame, we see that 26,397 flights departed New York City in 2013.

## LC4.30

Why should pie charts be avoided and replaced by barplots?

**Discussion**: 

Pie charts rely on us being able to tell the different between different slices of the pie in terms of the percentage of area covered.  Our eyes and brains are not naturally good at this task.  We are extremely good at determining the height of bars and comparing them across different groups as a barplot shows.

## LC4.33

What can you say, if anything, about the relationship between airline and airport in NYC in 2013 in regards to the number of departing flights?

**Discussion**: 

Certain airlines tend to depart from one airport far more than the other two airports.  For example, ExpressJet Airlines has nearly all of its flights from Newark.

## LC4.34

Why might the side-by-side barplot be preferable to a stacked barplot in this case?

**Discussion**: 

It is harder to compare the frequency of flights for LaGuardia or JFK when looking at a stacked barplot since they are already stacked on top of Newark.

## LC4.36

Why is the faceted barplot preferred to the side-by-side and stacked barplots in this case?

**Discussion**: 

A faceted barplot allows for comparisons to be made more easily across both airports AND airlines.

## LC4.37

What information about the different carriers at different airports is more easily seen in the faceted barplot?

**Discussion**: 

The faceted barplot provides an easier way to see the airlines that have no departing flights and also many flights from each airport.

***

# Problem Set 6 {#ps6-key}

Learning Checks [LC5.1-LC5.2, LC5.4-LC5.5, LC5.7, LC5.9-LC5.14](https://ismayc.github.io/moderndiver-book/5-manip.html)

## LC5.1

What’s another way using `!` we could filter only the rows that are not going to Burlington, VT nor Seattle, WA in the `flights` data frame? Test this out using the code above.

**Discussion**:

## LC5.2

Say a doctor is studying the effect of smoking on lung cancer of a large number of patients who have records measured at five year intervals. He notices that a large number of patients have missing data points because the patient has died, so he chooses to ignore these patients in his analysis. What is wrong with this doctor’s approach?

**Discussion**:

## LC5.4

Why doesn’t the following code work? You may want to run the code line by line instead of all at once. In other words, run `summary_temp <- weather %>% summarize(mean = mean(temp, na.rm = TRUE))` first.

```{r eval=FALSE}
summary_temp <- weather %>%   
  summarize(mean = mean(temp, na.rm = TRUE)) %>% 
  summarize(std_dev = sd(temp, na.rm = TRUE))
```


**Discussion**:

## LC5.5

Recall from Chapter [4](https://ismayc.github.io/moderndiver-book/4-viz.html) when we looked at plots of temperatures by months in NYC. What does the standard deviation column in the `summary_monthly_temp` data frame tell us about temperatures in New York City throughout the year?

**Discussion**:

## LC5.7

Recreate `by_monthly_origin`, but instead of grouping via `group_by(origin, month)`, group variables in a different order `group_by(month, origin)`. What differs in the resulting data set?

**Discussion**:

## LC5.9

How does the `filter` operation differ from a `group_by` followed by a `summarize`?

**Discussion**:

## LC5.10

What do positive values of the `gain` variable in `flights` correspond to? What about negative values? And what about a zero value?

**Discussion**:

## LC5.11

Could we create the `dep_delay` and `arr_delay` columns by simply subtracting `dep_time` from `sched_dep_time` and similarly for arrivals? Try the code out and explain any differences between the result and what actually appears in `flights`.

**Discussion**:

## LC5.12

What can we say about the distribution of `gain`? Describe it in a few sentences using the plot and the `gain_summary` data frame values.

**Discussion**:

## LC5.13

Looking at Figure [5.7](https://ismayc.github.io/moderndiver-book/5-manip.html#fig:reldiagram), when joining `flights` and `weather` (or, in other words, matching the hourly weather values with each flight), why do we need to join by all of `year`, `month`, `day`, `hour`, and `origin`, and not just `hour`?

**Discussion**:

## LC5.14

What surprises you about the top 10 destinations from NYC in 2013?

**Discussion**: